Revisiting divisible, injective and flabby abelian groups from a constructive point of view

### 2022-09-27 slides-fischbachau2022.tex
typo
Title slide
Introduction to ¬¬-embedding

\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd,wasysym}
\usepackage{bussproofs}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

@@

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.00cm}\hfill\mbox{\includegraphics[width=0.2\textwidth]{turing-machine}\;\;}\end{minipage}}
\begin{frame}{Exploring the realizability model}
  \medskip
  \medskip
  \medskip
  \begin{tabular}{@{\!\!\!\!\!\!\!}l@{\,}llp{1.8cm}}
    \toprule
    & Statement & classical? & realizable? \\
    \midrule
    \normalnumber{1} & Every number is prime or not prime. & \ccmark{}
    (trivially) & \ccmark \\
    \normalnumber{2} & After every number there is a prime. & \ccmark & \ccmark \\
    \normalnumber{3} & Every map $\NN \to \NN$ has a zero or not. & \ccmark{} (trivially) & \cxmark \\
    \normalnumber{4} & Every map $\NN \to \NN$ is computable. & \cxmark &
    \qswitch{4}{5}{\ccmark}\only<1-4>{\,} \visible<5->{(trivially)} \\
    \normalnumber{5} & Every map $\RR \to \RR$ is continuous. & \cxmark &
    \qswitch{5}{6}{\ccmark} \\
    \normalnumber{6} & Markov's principle holds. & \cmark{} (trivially) &
    \qswitch{6}{7}{\ccmark{} (if MP)} \\
    \normalnumber{7} & Countable choice holds. & \xmark &
    \qswitch{7}{8}{\ccmark{} (always!)} \\
    \normalnumber{8} & Heyting arithmetic is categorical. & \xmark &
    \qswitch{8}{9}{\ccmark{} (if MP)} \\
    \bottomrule
  \end{tabular}
  \medskip

@@

    \visible<4->{``Every function $\NN \to \NN$ has a zero or not'' is realized by an
    infinite search.}
    \medskip

@@

>>>>

\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd,wasysym}
\usepackage{bussproofs}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage{bbm}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

@@

\section{Realizability theory}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{turing-machine}
  \medskip

  \large Lecture I: \\
  \hil{Realizability theory}

  \normalsize

  \emph{for extracting programs from constructive proofs}
\end{frame}

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.00cm}\hfill\mbox{\includegraphics[width=0.2\textwidth]{turing-machine}\;\;}\end{minipage}}
\begin{frame}{Exploring the realizability model}
  \medskip
  \medskip
  \medskip
  \begin{tabular}{@{\!\!\!\!\!\!\!}l@{\,}llp{1.8cm}}
    \toprule
    & Statement & classical? & realizable? \\
    \midrule
    \normalnumber{1} & Every number is prime or not prime. & \ccmark{}
    (trivially) & \ccmark \\
    \normalnumber{2} & After every number there is a prime. & \ccmark & \ccmark \\
    \normalnumber{3} & Every map $\NN \to \NN$ has a zero or not. & \ccmark{} (trivially) & \cxmark \\
    \normalnumber{4} & Every map $\NN \to \NN$ is computable. & \cxmark &
    \qswitch{4}{5}{\ccmark}\only<1-4>{\,} \visible<5->{(trivially)} \\
    \normalnumber{5} & Every map $\RR \to \RR$ is continuous. & \cxmark &
    \qswitch{5}{6}{\ccmark} \\
    \normalnumber{6} & Markov's principle holds. & \ccmark{} (trivially) &
    \qswitch{6}{7}{\ccmark{} (if MP)} \\
    \normalnumber{7} & Countable choice holds. & \ccmark &
    \qswitch{7}{8}{\ccmark{} (always!)} \\
    \normalnumber{8} & Heyting arithmetic is categorical. & \cxmark &
    \qswitch{8}{9}{\ccmark{} (if MP)} \\
    \bottomrule
  \end{tabular}
  \medskip

@@

    \visible<4->{``Every function $\NN \to \NN$ has a zero or not'' is realized by
    infinite search.}
    \medskip

@@

\section{Proof transformations}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{philosophers-stone}
  \medskip

  \large Lecture II: \\
  \hil{Proof transformations}

  \normalsize

  \emph{for extracting constructive proofs from classical proofs}
\end{frame}

\begin{frame}{The double-negation embedding}
  \textbf{Thm.} Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
  in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.

  \textbf{Thm.} Every infinite binary sequence~$\alpha : \NN \to \mathbbm{2}$
  contains repeating entries.

  {\emph{Proof.} By the \bad{infinite box principle}, which follows
  from~\badbox{\textsc{lem}}, the sequence contains infinitely many zeros or
  infinitely many ones. In either case the claim follows.\par}

  How to extract \good{constructive proofs} from these
  \bad{classical proofs}?
  \pause

  \grayline

  \textbf{Thm.} Let~$X$ be an inhabited set of natural numbers. Then~$X$
  contains a minimal element.

  {\emph{Proof.} By~\badbox{\textsc{lem}}, \ldots\par}
\end{frame}
